---
title: "Exploratory Data Analysis"
author: "Logan Wu"
date: "7 January 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=T, cache=T)
library(tidyverse)
library(data.table)
library(sf)
library(leaflet)
library(spdep) #easily create weight matrices
library(starma)
library(forecast)
library(ggcorrplot)

# read in a manageable portion of the dataset for now
raw.ts <- fread("data/Pedestrian_volume__updated_monthly_.csv") %>% head(25000)
raw.ts[,Date_Time := as.POSIXct(Date_Time, format="%m/%d/%Y %I:%M:%S %p")]

raw.geo <- read.csv("data/Pedestrian_sensor_locations.csv")
```

## Process TS data

```{r}
data.ts = raw.ts %>%
  select(-ID)

ggplot(data.ts %>% filter(Sensor_ID %in% seq(6)), aes(x=Date_Time, y=Hourly_Counts, color=Sensor_ID)) +
  geom_line(alpha=0.5) +
  facet_grid(Sensor_Name~.) +
  guides(color=F)
```

Data is incomplete for some sensors. Possible reasons include maintenance.

```{r}
ggplot(data.ts, aes(x=Hourly_Counts)) +
  geom_density(alpha=0.5, fill="red", color=NA)

ggplot(data.ts, aes(x=log(Hourly_Counts+1))) +
  geom_density(alpha=0.5, fill="blue", color=NA)
```


```{r}
data.ts2 <- data.ts %>%
  select(Date_Time, Sensor_Name, Hourly_Counts) %>%
  spread(key=Sensor_Name, value=Hourly_Counts) %>%
  select(-Date_Time) %>%
  as.ts %>%
  na.contiguous # analysis does not include missing data
  # stcenter# %>%

ggcorrplot(cor(data.ts2[,1:16]), show.legend=F, type="lower", lab=T, lab_size=1.5)
```

Spatial correlations have varying levels -- factors other than time of day/year lead to correlation.

```{r}

```

## Experiment with models

Fit a simple time series without any cross-correlation

```{r}
result <- lapply(data.ts2, function(x) forecast(auto.arima(x)))
plot(result[[1]])
```

## Multiple ARIMA

```{r}
# marima.result = marima(data.ts2, ar.pattern=2)
```

Ignore this: example from website

```{r}
# Create a 5x5 regular grid which will be our lattice
sites <- matrix(0, 25, 2)
for (i in 1:5) {
	for (j in 1:5)
		sites[(i-1)*5 + j, ] <- c(i, j) - .5
}
plot(sites)

# Create a uniform first order neighbourhood
knb <- dnearneigh(sites, 0, 1)
plot(knb, sites)

# Lag the neighbourhood to create other order matrices
knb <- nblag(knb, 4)
klist <- list(order0=diag(25),
           order1=nb2mat(knb[[1]]),
           order2=nb2mat(knb[[2]]),
           order3=nb2mat(knb[[3]]),
           order4=nb2mat(knb[[4]]))
		   
# Simulate a STARMA(2;1) process
eps <- matrix(rnorm(200*25), 200, 25)
star <- eps
for (t in 3:200) {
	star[t,] <- (.4*klist[[1]] + .25*klist[[2]]) %*% star[t-1,] + 
	  (.25*klist[[1]]) %*% star[t-2,] + (-.3*klist[[2]]) %*% eps[t-1,] + 
	  eps[t, ]
}

star <- star[101:200,]	# Remove first observations
star <- stcenter(star)	# Center and scale the dataset
										   
# Identify the process
stacf(star, klist)
stpacf(star, klist)

# Estimate the process
ar <- matrix(c(1, 1, 1, 0), 2, 2)
ma <- matrix(c(0, 1), 1, 2)
model <- starma(star, klist, ar, ma)
model
summary(model)

# Diagnose the process
stcor.test(model$residuals, klist, fitdf=4)
stacf(model$residuals, klist)
stpacf(model$residuals, klist)

# starma(matrix.ts, c(diag(18)), ar=0, ma=0)
```

## Process geographic data

```{r}
data.geo = raw.geo %>%
  mutate(Latitude = as.numeric(Latitude),
         Longitude = as.numeric(Longitude)) %>%
  # st_as_sf(coords=c("Latitude", "Longitude")) %>%
  select(-Status, -Upload.Date, -Geometry, -Location.Type)

leaflet() %>%
  addProviderTiles(providers$CartoDB.Positron) %>%
  addCircleMarkers(data.geo$Longitude, data.geo$Latitude,
                   label=data.geo$Sensor.Description,
                   weight=0, radius=4, fillOpacity=0.5)
```

## Estimate neighborhoods based on the example

```{r}
sites = data.geo %>% select(Latitude, Longitude) %>% as.matrix
knb <- dnearneigh(sites, 0, 0.006)
plot(knb, sites)
```

```{r, eval=F}
# Lag the neighbourhood to create other order matrices
knb <- nblag(knb, 3)
klist <- list(order0=diag(58),
           order1=nb2mat(knb[[1]]),
           order2=nb2mat(knb[[2]]),
           order3=nb2mat(knb[[3]]))

# Identify the process
stacf(data.ts2, klist)
stpacf(data.ts2, klist)
```